Formal Verification Framework Based on Symbolic Execution for Smart Contract

### 2023-12-22 papers/static-checker/ecoop2024/ecoop2024.tex
write USD part

\begin{lstlisting}[float,captionpos=b,caption={Smart contract property specification},label={lst:smart-contract-property-specification},numbers=left]
mcontract Name = spec
code := {c}
input := t
output := t'
pre-condition := cd
post-condition := cd'
\end{lstlisting}
Listing \ref{lst:smart-contract-property-specification} shows the syntax of our property specification language. Here, \KMCONTRACT, \KWITH, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are keywords. The \KNAME\ is replaced by the contract name. In the input and output fields, the input and output patterns are specified. Since Michelson operates on a stack, we assume that the stack has only one element, as specified in the input pattern. Similarly, the output stack is assumed to be a stack with only one element, which is specified in the output field. As we have mentioned before, the output should be a pair consisting of an operation list and a storage. In our syntax, for verification purpose, we omit the operation list and then the output is represented by the patern of the value in the storage after the execution is finished. The input and output patterns represent Michelson terms, which can be either concrete values or symbolic values, including variables. When a variable appears for the first time, it needs to be accompanied by its corresponding type declaration. However, subsequent mentions can omit the type declaration. Formally, the pre and post-condition instances are formulas in first-order logic, and they can be handled by an SMT solver, namely the Z3 solver in our framework. The detail of the language grammar is described in in Figure \ref{}. Formulas that are used in assertions or added to solvers are terms of Boolean sort. Otherwise, terms of Boolean and non-Boolean sort may be mixed in any combination where sorts match up. Universal and existential quantifiers bind variables to the scope of the quantified formula.
\begin{lstlisting}[float,captionpos=b,caption={The specification of Add contracr},label={lst:add-contract-specification},numbers=left]
mcontract Add = spec
code := {UNPAIR; COMPARE; GE; 
        IF {ADD}{PUSH string 'Unexpected Pair'; FAILWITH}}
input := Pair (x : int) (y : int)
output := (z : int)
pre-condition := (x >= 5) | (y > 3)
post-condition := (z >= 8)
\end{lstlisting}
Let's consider the  example of the property specification for the contract named Add as in Listing \ref{lst:add-contract-specification}. Here, the program takes as input a stack with only one element, which is a pair of two integers, symbolically specified in the Input field. It then unpairs the input pair to obtain two elements as two integers \lstinline/x/ and \lstinline/y/.  It then proceeds to destructure the input pair, yielding two integers. Subsequently, the program performs a comparison between these two integers. If  \lstinline/x/ is greater than or equal to \lstinline/y/, the program constructs a new stack with a solitary element at the top representing the sum of  \lstinline/x/ and \lstinline/y/, designated as \lstinline/z/. Otherwise, the execution terminates with an error message containing the content 'Unexpected pair.' The specification continues by defining pre and post conditions, where the precondition states the assumption that the integer  \lstinline/x/ is greater than or equal to 5 and \lstinline/y/ is greater than or equal to 3. Here, the symbol  \lstinline/|/ is the conjuration. The postcondition asserts that the result sum \lstinline/z/ is greater than or equal to 8.

@@

There are several properties that we want to guarantee in this contract. Let's take a look at the specification for the auction contract. For the \lstinline/bid/ entrypoint, the input stack is the one with only one element, which is a pair consisting of the parameter and the current storage. Here, the parameter is \lstinline/Unit/ of type \lstinline/unit/. The current storage is represented symbolically as a nested pair, where the first element is the \lstinline/auction_open/ variable of type \lstinline/bool/, and the second element is another pair consisting of the \lstinline/highest_bidder/ of type \lstinline/address/ as the first element and the \lstinline/contract_owner/ variable, also of type \lstinline/address/, as the second element. The output term should maintain the required pattern of the storage, where the values may be changed, and these changes are recorded in other symbolic variables. In this specification, we add the prefix \lstinline/post/ to the old variable names to represent the new variables corresponding to the old ones.

@@

\subsubsection{Life-cycle check}
\section {Case Studies}
\subsection{USDtz}

\subsection{Kolibri Oracle Contract}
\section {Related Work}
\section {Conclusion}
\section {Appendix}
\end{document}

>>>>

\begin{lstlisting}[float,captionpos=b,caption={Smart contract property specification},label={lst:smart-contract-property-specification},numbers=left]
mcontract Name = spec
code := {c}
input := t
output := t'
pre-condition := cd
post-condition := cd'
\end{lstlisting}
Listing \ref{lst:smart-contract-property-specification} shows the syntax of our property specification language. Here, \KMCONTRACT, \KWITH, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are keywords. The \KNAME\ is replaced by the contract name. In the input and output fields, the input and output patterns are specified. Since Michelson operates on a stack, we assume that the stack has only one element, as specified in the input pattern. Similarly, the output stack is assumed to be a stack with only one element, which is specified in the output field. Our syntax allows to specify incomplete and complete smart contract code. For the complete smart contrat code, as we have mentioned before, the output should be a pair consisting of an operation list and a storage.In our syntax, for verification purposes, we exclude the operation list, and the output is solely represented by the pattern of the value in the storage after the execution concludes. In fact, for a complete smart contract code, in some cases, the output pattern can be omitted, given that the input pattern must be a pair of parameters and storage. This implies that the storage pattern is part of the input pattern. In such cases, the prefix 'post' is used before the variables in the input term to indicate the corresponding storage value after the execution.  The input and output patterns represent Michelson terms, which can be either concrete values or symbolic values, including variables. When a variable appears for the first time, it needs to be accompanied by its corresponding type declaration. However, subsequent mentions can omit the type declaration. Formally, the pre and post-condition instances are formulas in first-order logic, and they can be handled by an SMT solver, namely the Z3 solver in our framework. The detail of the language grammar is described in in Figure \ref{}. Formulas that are used in assertions or added to solvers are terms of Boolean sort. Otherwise, terms of Boolean and non-Boolean sort may be mixed in any combination where sorts match up. Universal and existential quantifiers bind variables to the scope of the quantified formula.
\begin{lstlisting}[float,captionpos=b,caption={The specification of Add contracr},label={lst:add-contract-specification},numbers=left]
mcontract Add = spec
code := {UNPAIR; COMPARE; GE; 
        IF {ADD}{PUSH string 'Unexpected Pair'; FAILWITH}}
input := Pair (x : int) (y : int)
output := (z : int)
pre-condition := (x >= 5) | (y > 3)
post-condition := (z >= 8)
\end{lstlisting}
Let's consider the  example of the property specification for the contract named Add as in Listing \ref{lst:add-contract-specification}. Here, the program takes as input a stack with only one element, which is a pair of two integers, symbolically specified in the Input field. It then unpairs the input pair to obtain two elements as two integers \lstinline/x/ and \lstinline/y/.  It then proceeds to destructure the input pair, yielding two integers. Subsequently, the program performs a comparison between these two integers. If  \lstinline/x/ is greater than or equal to \lstinline/y/, the program constructs a new stack with a solitary element at the top representing the sum of  \lstinline/x/ and \lstinline/y/, designated as \lstinline/z/. Otherwise, the execution terminates with an error message containing the content 'Unexpected pair.' The specification continues by defining pre and post conditions, where the precondition states the assumption that the integer  \lstinline/x/ is greater than or equal to 5 and \lstinline/y/ is greater than or equal to 3. Here, the symbol  \lstinline/|/ is the conjuration. The postcondition asserts that the result sum \lstinline/z/ is greater than or equal to 8.

@@

There are several properties that we want to guarantee in this contract. Let's take a look at the specification for the auction contract. For the \lstinline/bid/ entrypoint, the input stack is the one with only one element, which is a pair consisting of the parameter and the current storage. Here, the parameter is \lstinline/Unit/ of type \lstinline/unit/. The current storage is represented symbolically as a nested pair, where the first element is the \lstinline/auction_open/ variable of type \lstinline/bool/, and the second element is another pair consisting of the \lstinline/highest_bidder/ of type \lstinline/address/ as the first element and the \lstinline/contract_owner/ variable, also of type \lstinline/address/, as the second element. The output term should maintain the required pattern of the storage, where the values may be changed, and these changes are recorded in other symbolic variables. In this specification, we add the prefix \lstinline/post/ to the old variable names to represent the new variables corresponding to the old ones.Note that the output pattern here could be safely removed, and the tool will automatically generate the output pattern to match the one specified here.

@@

\subsubsection{Life-cycle check}
\section {Case Studies}
\subsection{USDtz}
In a blockchain ecosystem, tokens represent digital assets that can be transferred between accounts. These tokens are securely stored and verified within the blockchain network. There are various types of tokens, including native tokens (cryptocurrencies) like Bitcoin, Ether, Tezos, and others. Additionally, tokens can be created through smart contracts on an existing blockchain. Two common categories are: (1) Fungible Tokens (FT), which are interchangeable and (2) Non-Fungible Tokens (NFTs), which represent unique digital assets, and each unit is distinct. Stablecoins, falling under fungible tokens, are designed to maintain a stable value by tracking the value of a fiat currency like USD or EUR.

While you can create tokens that behave in any way that you want them to behave, implementing tokens that adhere to established standards is advisable for interoperability and compatibility across different platforms. A Token standard defines a set of rules and an interface that smart contracts must follow to ensure consistency and compatibility with decentralized exchanges, wallets, and other tools within the ecosystem. For example, ERC-20 is a stoken tandard used in Ethereum.

Tezos blockchain has its own set of standards known as Financial Applications (FA). Two notable standards are: (1) FA 1.2, designed for fungible tokens, and (2) FA 2, a more versatile standard accommodating both fungible and non-fungible tokens (NFTs).  The standard outlines a smart contract that functions as a ledger, associating identities with balances. This ledger is designed to facilitate token transfer operations and enable approvals for spending tokens from other accounts. Any contract aiming to implement the standard must incorporate the functions specified in Listing \ref{lst:fa12-interface} such as transfer (transacting between users), approve (allowing one user to send tokens from another user's address), allowance (browsing the amount of tokens available for an account to sending from another address), user balance (checking the balance of a user), and current supply (checking the total supply of tokens across all addresses). Further elaboration on the details of FA 1.2 can be found in \ref{}, and we will delve into these functions in subsequent sections.

\begin{lstlisting}[float,captionpos=b,caption={FA 1.2 interface},label={lst:fa12-interface},numbers=left]
(address :from, (address :to, nat :value)) %transfer
(address :spender, nat :value) %approve
(view (address :owner, address :spender) nat) %getAllowance
(view (address :owner) nat) %getBalance
(view unit nat) %getTotalSupply
\end{lstlisting}

In this section, we verify USDtz \ref{}, an implementation of the FA 1.2 standard. Our goal is to ensure that the implementation aligns with the standard, verifying its correctness. USDtz is a stablecoin and one of the first and longest-running coins on the Tezos blockchain, launched in September 2020. The smart contract is located at the address KT1LN4LPSqTMS7Sd2CJw4bbDGRkMv2t68Fy9, and the code can be found on the Tezos blockchain \ref{}. USDtz adheres to the FA token standard, and the initial deployment of the USDtz smart contract follows the FA 1.2 standard.
The implementation consists of approximately 1000 instructions. It contains all of five entrypoints required in the FA 1.2 interface and four others.  We will verify this smart contract for each of its entrypoints.  Let's first examine how the approvable ledger is implemented in the USDtz smart contract. The ledger is a substantial map that links user addresses to a pair consisting of their balance and a map associating addresses with the amount of tokens allowed to be transferred from the user's account. The storage structure of the USDtz smart contract is illustrated in Listing \ref{lst:storage-udstz-contract}.

\begin{lstlisting}[float,captionpos=b,caption={Storage of the USDtz smart contract},label={lst:storage-udstz-contract},numbers=left]
  ( pair
    ( big_map %ledger
      ( address :user )
      ( pair ( nat :balance ) ( map :approvals
        ( address :spender )
        ( nat :value ) ) ) )
    ( pair ( address %admin ) 
           ( pair ( bool %paused ) ( nat %totalSupply ) ) ) ) 
\end{lstlisting}

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the \lstinline/transfer/ entrypoint credits the account associated with the address provided in the \lstinline/to/ parameter, while debiting the account corresponding to the \lstinline/from/ parameter with the specified \lstinline/value/ amount. The FA 1.2 standard mandates that this call must fail under the following two conditions: (1) In the event of insufficient balance, this entrypoint MUST fail with the error \lstinline/NotEnoughBalance/, indicating insufficient funds in the sender's account to execute the transfer. The error MUST contain a \lstinline/pair (nat :value) (nat :present)/ pair, where \lstinline/value/ is the requested amount of tokens, and \lstinline/present/ is the available amount. (2) If the sender's account has no permission to withdraw the \lstinline/value/ amount of funds, the call will fail with the \lstinline/NotEnoughAllowance/ error, which also contains the same pair message as the first failure scenario. To verify these requirements, we unize the failwith condition of the tool and the result return as in Listing \ref{}.

\subsection{Kolibri Oracle Contract}
\section {Conclusion}
\section {Appendix}
\end{document}
